---
title: "eds221-day3"
format: html
editor: visual
---

## Conditional Statement

```{r}
#This is a plain if statement 
temperature <- 95

if(temperature > 97) {
  print("temperature warning")
}

#if else:

if(temperature > 97) {
  print("temperature warning")
} else
  { print("no temperature warning")
}
```

Create an if else statement for blue wale lengths. Create a variable blue whale, which has a value of 40 meters. Add an if else statement that returns "GIANT" if the length exceeds 35, "MEDIUM" if the length is between 20 and 35, or "SMALL" if the length \<= 20.

```{r}
blue_whale <- 40

if(blue_whale > 35) {
  print("GIANT")
} else if (blue_whale >= 20 & blue_whale <= 35)
  { print ("MEDIUM")
} else if (blue_whale < 20) {
  print("SMALL")}
```

## Switch statement

```{r}
#This is one option to writing better if else statements, create a new value 
housing_type <- "single"

switch(housing_type, 
       "single" = print("400"),
       "double" = print("750"),
       "whole" = print("1100"))

if(housing_type == "whole") {
  print("1100")
} else if (housing_type == "double") {
  print("750")
} else if (housing_type == "single") {
  print ("400")
}
       
```

## For loops

```{r}
height <- c(58, 61, 76, 68, 85, 75)

for (i in seq_along(height)) {
  half_height = height[i] / 2
  print(half_height)
}

##For each value in the height vector, retun the sum of the value + the one right after it

for (i in seq_along(height)) {
  plusone = (height[i] + height[i + 1])
  print(plusone)
}

```

```{r}
##Write a for loop that calculates the value of height for each element divided by the mean of the entire height vector 

for (i in seq_along(height)) {
 mean = height[i] / mean(height)
 print(mean)
}
```
